### Read the csv file

```{r}
setwd('C:/Users/daala/Desktop/R course/Reproducible research/Week 2')
activity <- read.csv("activity.csv")
activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity, weekday)
```



### Ques 2
###Histogram of the total number of steps taken each day

```{r}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm =TRUE))
names(activity_total_steps) <- c("date","steps")
hist(activity_total_steps$steps, main = "total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000,by = 2500))
```

### Ques 3
### Mean and median of the total steps taken
```{r}
rmean <- mean(activity_total_steps$steps)

rmeadian <- median(activity_total_steps$steps)
```

### what is the average daily activity patterns?
```{r}
steps_by_interval <- aggregate(steps~interval, activity, mean)
plot(steps_by_interval$interval, steps_by_interval$steps, xlab="Interval", ylab = "Number of steps", main = "Avg numer of steps per day")

# 5-minute interval on avg across all the days in the data set
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
```

### Impute missing values. Compare imputed to non-imputed data
Missing values were imputed bu inserting the avg for each interval
```{r}
not_complete <- sum(!complete.cases(activity))
imputed_data <- transform(activity, steps = ifelse(is.na(activity$steps), steps_by_interval$steps[match(activity$interval,steps_by_interval$interval)], activity$steps))
```

Zeroes to be imputed for 10-01-2012 because it was the first day
```{r}
imputed_data[as.character(imputed_data$date) == 2012-10-01, 1] <- 0
```

Recount total steps by day and create histogram
```{r}
steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("total steps Each day"), col="blue",xlab="Number of steps")

#create histograms to show diff

hist(steps_by_day_i$steps, main = paste("total steps each day"), col="red", xlab="Number of steps", add=T)
legend("topright", c("Imputed","Non-Imputed"), col=c("blue","red"), lwd=10)
```

Mean and median of the data
```{r}
rmean_i <- mean(steps_by_day_i$steps)
rmedian_i <- median(steps_by_day_i$steps)
```

Difference between imputed and non-imputed data
```{r}
mean_diff <- rmean_i - rmean
median_diff <- rmedian_i - rmedian
```

Total difference
```{r}
total_diff <- sum(steps_by_day_i$steps) - sum(steps_by_day_i$steps)
```


### Is there any difference between activities in weekdays and weekends
```{r}
weekdays <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays),"Weekday","Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)
xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main = "Avg Steps per day by interval", xlab="Interval",ylab="Steps",layout=c(1,2), type="1")
```

